---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r}
        setwd("/Users/krishna/Desktop/RepData_PeerAssessment1-master")
        rm(list=ls())
        library(lattice)
        dat <- read.csv("activity.csv")
        dat$date <- as.Date(dat$date) # convert dates from factor to datetime
```

## What is mean total number of steps taken per day?

```{r}
        steps <- aggregate(x=dat$steps,by = list(dat$date),FUN = "sum")$x # steps col
        hist(steps)
        cat("The mean total number of steps taken per day is:",mean(na.omit(steps)),"\n")
        cat("The median total number of steps taken per day is:",median(na.omit(steps)),"\n")
```



## What is the average daily activity pattern?

```{r}
        daily <-aggregate(x=dat$steps,by = list(dat$interval),FUN = "mean",na.rm ="TRUE")
        names(daily)[1] <-paste("interval") # give columns descriptive names
        names(daily)[2] <-paste("meanval")
        plot(daily$meanval,type="l", 
             xlab="5-minute interval #",
             ylab="average number of steps taken",
             main="Avg number of steps taken VS 5-minute interval #")
        cat("The 5-minute interval that, on average, contains the maximum number of steps:",which(daily$meanval == max(daily$meanval)),"\n")
```


## Imputing missing values

```{r}
        cat("The total number of missing values in the dataset:",sum(is.na(dat$steps)),"\n")
```
### Code strategy for imputing missing data
We'll replace NAs with the mean for each 5-minute interval averaged over all days in the dataset, and create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
        newdat <- dat
        newdat$steps[is.na(newdat$steps)] <- daily[match(newdat$interval[is.na(newdat$steps)],daily$interval),"meanval"]
        steps <- aggregate(x=newdat$steps,by = list(newdat$date),FUN = "sum")$x # steps col
        hist(steps,
             xlab="Number of Steps taken each day",
             ylab="Frequency of occurence",
             main="Histogram of the Total Number of Steps Taken Each Day")
        cat("The mean total number of steps taken each day (with imputed data) is:",mean(steps),"\n")
        cat("The median total number of steps taken each day (with imputed data) is:",median(steps),"\n")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
        newdat[,"typeday"] <- factor((weekdays(dat$date) == "Saturday" | weekdays(dat$date) == "Sunday"), labels=c("weekday", "weekend"))
        avgsteps <- aggregate(newdat$steps~newdat$interval+newdat$typeday,FUN=mean)
        names(avgsteps)[1] <-paste("interval") # give columns descriptive names
        names(avgsteps)[2] <-paste("typeday")
        names(avgsteps)[3] <-paste("meanval")
        xyplot(avgsteps$meanval~avgsteps$interval | avgsteps$typeday, 
               layout = c(1,2),type="l",
               xlab="Interval",
               ylab="Number of Steps")
```
Yes there are differences in activity patterns between weekdays and weekends, as the graphs above indicate.
